---
title: "Gov 50 Final Project"
author: "Alicia Campbell"
description: "Measuring the Relationship between Party Religiousity and Female Membership in Lebanon"
output:
  distill::distill_article:
    self_contained: false
--- 



## Project thoughts (Milestone One)

I am interested in using the data set "Party Variation in Religiosity and Womens Leadership, Lebanon Dataset" to measure the relationship between party religiosity and share of women in party membership (perc) in Lebanon. Based on the results, I will examine the extent that party religiosity is associated with female membership, or lack thereof.


## Project Proposal (Milestone Two)

Research Question: Does party religiosity impact female party membership in Lebanon? How do other variables such as year of party origin play a role in level of religiosity and share of female membership ?

Drawing on the “Party Variation in Religiosity and Women's Leadership, Lebanon Dataset," I aim to examine the extent to which the religiosity of key political parties in Lebanon impacts the share of female membership in said party. I hypothesize that the share of female membership decreases as party religiosity increases. As is typical, I expect more religious parties to take more conservative stances with regards to the role of women in the political realm, which is often associated with an adherence to gender roles wherein women are seen as better suited to carry out more domestic-oriented, than public-facing tasks. Conversely, I hypothesize that less religious parties in Lebanon will exhibit higher shares of female membership, given the more liberal and open ideology often associated with non-religious groups. My sample includes responses regarding religion and female membership from 18 different political parties in Lebanon of different levels of religiosity. My explanatory variable is level of party religiosity, with possible answers ranging from “Areligious and Secular” (coded as 5) to “Extremely Religious” (coded as 1).The full ranking of choices are as follows: 1) Extremely Religious, 2) Conservative but not extremist religiosity 3) tolerant religiosity 4) confessional but civil secularism 5) A religious and Secular. My outcome variable is the share of female members in a given political party, represented as a percent.  For example, if the share of a female membership is 15, it would indicate that 15% of party members are female. According to the dataset summary, this data was collected “in person and via one-on-one interviews with party administrators, officials, and elites.” If I were to observe lower shares of female party membership for parties that are coded as having high levels of religiosity (likely coded as a 1 or 2), this would support my hypothesis. Conversely, if I were to observe the opposite effect, that female membership is higher for parties that exhibit higher levels of religiosity, this would provide evidence against my hypothesis. 


```{r}
library(tidyverse)
library(ggplot2)
library(scales)

#loading data
library(haven)
women_religion <- read_sav("Party Variation in Religiosity and Womens Leadership Lebanon Dataset.SAV")

#grouping by level of religiosity and calculating the mean female party membership for each group
fem_membership_mean <- women_religion %>%
  group_by(SECLRSM) %>%
 summarize(
   fem_avg = mean(F_MEM, na.rm = TRUE))

fem_membership_mean 

#configuring x axis labels
xlabels <- c("Areligious and secular",
  "Confessional but civil secularism",
  "Tolerant religiosity",
  "Conservative but not extremist religiosity",
  "Extremist religiosity")

wrapped_labels <- str_wrap(xlabels, width = 19)


#producing bar plot showing the relationship between level of party religiosity and share of female membership 

rel_fem_plot <- ggplot(fem_membership_mean, mapping = aes(x = SECLRSM, y = fem_avg)) + geom_col(fill = "steelblue") +
  labs(
    title = "Religiosity and Female Party Membership",
    x = "Level of Religiosity",
    y = "Share of Female Membership in Party"
  ) + 
   scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_x_continuous(
    breaks = as.numeric(fem_membership_mean$SECLRSM),
    labels =  wrapped_labels)


rel_fem_plot
```
The above bar plot shows the relationship between religiosity (on a scale from 1 - 5, 1 being the least religious) and share of female party membership (reflected as a percentage). From this graph we can see that the share of female membership generally decreases as religiosity increases. The greatest level of female membership is seen at level 2, dubbed "confessional but civil secularism" and the lowest is at level 5, dubbed "extremist religiosity." However, the exact extent of the relationship between these two variables is not entirely clear without further analysis.



```{r}
yr_fem_religion <- lm(F_MEM ~ SECLRSM, data = women_religion)
yr_fem_religion
```
Interpretation of regression: Religiosity on share of female party membership:

In this case, the intercept of 39.808 has no practical meaning since religiosity is will never be 0. The scale of religiosity only ranges from 1 - 5.

The SECLRSM coefficient (religiosity) is -3.634. This means that for a one level increase in religiosity, the share of female party membership is expected to decrease by 3.634 percentage points. Thus, it seems that on average, as religiosity increases, female party membership decreases by 3.634 percentage points.



```{r}
#single variable regression, year of origin on female membership
fem_membership_year_regression <- lm(F_MEM ~ YRORGN, data = women_religion)

fem_membership_year_regression
```
Interpretation of regression: Year of party origin on share of female party membership:

While the intercept is not a relevant measure here since the year will never be equal to 0, the coefficient of YRORGN (year of party origin) is 0.3143. This means that for each one-year increase in the year of party origin (as in one year closer to present time), the share of female party membership is expected to increase by 0.3143 percentage points. This suggests that on average, each year closer to present time that a party was established, the share of female membership will increase 0.3143 percentage points.



```{r}
#multivariable regression, year of origin AND level of religiosity on female membership

yr_fem_religion <- lm(F_MEM ~ YRORGN + SECLRSM, data = women_religion)
yr_fem_religion

```

Interpretation of regression: Religiosity AND year of party origin on share of female party membership:

The intercept has no practical meaning in this case since neither level of religiosity nor year of origin will ever be equal to 0. 

The coefficient of YRORGN is 0.2886. This means that, for each year closer to present time that a party was established, the the share of female party membership is expected to increase  0.2886 percentage points.

The coefficient for SECLRSM is -1.8041. This means that for each level increase in religiosity (SECLRSM), the share of female membership expected to decrease by 1.8041 percentage points.



```{r}
#visualization of relationship between level of party religiusity and female party membership
plot_religiosity <- ggplot(women_religion, aes(x = SECLRSM, y = F_MEM)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Relationship between Religiosity and Female Membership",
    x = "Religiosity",
    y = "Female Membership"
  )

plot_religiosity
```
```{r}
#visualization of relationship between year of party origin and female party membership

plot_year <- ggplot(women_religion, aes(x = YRORGN, y = F_MEM)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(
    title = "Relationship between Year of Origin and Female Membership",
    x = "Year of Origin",
    y = "Female Membership"
  )

plot_year
```
```{r}
plot_religiosity_year <- ggplot(women_religion, aes(x = YRORGN, y = SECLRSM)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Relationship between Religiosity and Year of Origin",
    x = "Year of Origin",
    y = "Religiosity"
  )

plot_religiosity_year
```









